---
title: "ranaAnalysisMarkdown"
author: "Maya Rayle"
date modified: "3/20/2024"
output:
  pdf_document: default
  html_document: default
  word_document: default
---

Loading and setting up the data, renaming some columns, adding a few columns, data wrangling etc. 

```{r setup, include=FALSE}
#Loading Packages
library(tidyverse)
library(ggplot2)
library(readr)
library(stringr)
library(lubridate)
library(RColorBrewer)
library(rstatix)
library(dplyr)
library(ggpubr)
library(gridExtra)

#set working directory
setwd("/Users/mayarayle/Documents/Ranavirus Project")

#loading in the data
MetaTransmit <- read_csv("Most Recently Updated Data/MetaTransmit.csv")

#---------------Data wrangling, cleaning up names etc

#getting rid of a couple extra lines that shouldn't be there
MetaTransmit <- MetaTransmit[-c(439:442),]

#renaming
names(MetaTransmit)[1] <- "Simple_ID"
names(MetaTransmit)[2] <- "Exposure_ID"
names(MetaTransmit)[3] <- "Death_timing"
names(MetaTransmit)[4] <- "Pre_exp_ID"
names(MetaTransmit)[5] <- "Exposure_type"
names(MetaTransmit)[6] <- "Exposure_start"
names(MetaTransmit)[7] <- "Exposure_duration"
names(MetaTransmit)[8] <- "Animal_type"
names(MetaTransmit)[9] <- "Death_date"
names(MetaTransmit)[12] <- "Erin_ID"
names(MetaTransmit)[13] <- "Well_1"
names(MetaTransmit)[14] <- "Well_2"
names(MetaTransmit)[15] <- "Well_3"
names(MetaTransmit)[16] <- "Status"
names(MetaTransmit)[17] <- "MeanCopy"
names(MetaTransmit)[18] <- "Well_1_Redo"
names(MetaTransmit)[19] <- "Well_2_Redo"
names(MetaTransmit)[20] <- "Well_3_Redo"
names(MetaTransmit)[21] <- "Status_Redo"
names(MetaTransmit)[22] <- "MeanCopy_Redo"

#making a few alterations
MetaTransmit <- MetaTransmit %>% 
  #Converting to Numerical, pulling out the "hrs"
  mutate(Exposure_duration_numerical = parse_number(Exposure_duration)) %>% 
  #making the start date a date 
  mutate(Exposure_start = as.Date(mdy(Exposure_start))) %>%
  #making the death date a date 
  mutate(Death_date = as.Date(mdy(Death_date))) %>%
  #setting mean copy final as a numeric value
  mutate(MeanCopyFinal = as.numeric(MeanCopyFinal)) %>% 
  #Log of MeanCopyFinal
  #there's an issue with the zeroes.... they're -inf. To solve this, I add + 1 to everything.
  #It's not a great solution but it's what I have right now...
  mutate(logMeanCopyFinalPlus1 = log10(MeanCopyFinal + 1)) %>% 
  #MeanCopyFinal + 1 (for the glm linear model)
  mutate(MeanCopyFinalPlus1 = MeanCopyFinal + 1) %>% 
  #create the column Trio ID (for pair matching) and modify it from Exposure ID
  mutate(Trio_ID = str_replace(Exposure_ID, pattern = ".S[HQ].", replacement = "."))

#Larval Treatment IDs -- adding a new column
MetaTransmit$Larval_condition <- ifelse(grepl("ALS", MetaTransmit$Exposure_ID), "ALS", NA)
MetaTransmit$Larval_condition[grepl("AHS", MetaTransmit$Exposure_ID)] = "AHS"
MetaTransmit$Larval_condition[grepl("EHS", MetaTransmit$Exposure_ID)] = "EHS"
MetaTransmit$Larval_condition[grepl("ELS", MetaTransmit$Exposure_ID)] = "ELS"

#Creating a single column for Status using Status and Status_Redo (from the redos)
MetaTransmit$Final_status[grepl("Neg", MetaTransmit$Status)] = "Neg"
MetaTransmit$Final_status[grepl("Pos", MetaTransmit$Status)] = "Pos"
MetaTransmit$Final_status[grepl("Neg", MetaTransmit$Status_Redo)] = "Neg"
MetaTransmit$Final_status[grepl("Pos", MetaTransmit$Status_Redo)] = "Pos"

#creating a binomial of infection status
MetaTransmit$Final_status_binomial[grepl("Pos", MetaTransmit$Final_status)] = "1"
MetaTransmit$Final_status_binomial[grepl("Neg", MetaTransmit$Final_status)] = "0"

#doing some factoring: 
MetaTransmit <- MetaTransmit %>% 
  #factor exposure duration in order
  mutate(Exposure_duration = factor(Exposure_duration, 
                                     levels = c("1hr", "6hr", "12hr", "24hr", "48hr"))) %>% 
  #factor larval condition in order
  mutate(Larval_condition = factor(Larval_condition, 
                                    levels = c("AHS", "ALS", "EHS", "ELS"))) %>% 

  #factor animal type in order
  mutate(Animal_type = factor(Animal_type, 
                               levels = c("Naive", "Infected","Control (pseudo-naive)", 
                                          "Control (pseudo-infected)")))


#there were a few trio IDs where the animal was paired with a different one than the id number. So I'm fixing these manually. 
#In Shared (Cohabitation), ELS.24.4 virus (MR119) was paired with ELS.24.6 naïve (MR326), and ELS.24.6 virus (MR404) was paired with ELS.24.4 naïve (MR324). 
#Even though the sequential naives were unaffected, I still need to make sure the pairs are the same. So I am just changing the naive shared animals to the opposite trio ID. 
  #ELS.24.6 naïve (MR326) --> ELS.24.4
  #ELS.24.4 naïve (MR324) --> ELS.24.6

# Define the conditions and replacement values
conditions <- c("MR326", "MR324")
replacement_values <- c("ELS.24.4", "ELS.24.6")
 
# Use replace() to replace the names in the 'Names' column
MetaTransmit$Trio_ID <- replace(MetaTransmit$Trio_ID, MetaTransmit$Simple_ID %in% conditions, replacement_values)

#---------------------------------------------------------------------------------------------------------------------
#Data Checks - making sure that everything was entered correctly. 
# Add a new column to store the result
MetaTransmit$ContainsNumeric <- FALSE
MetaTransmit$MatchesExposureType <- FALSE

# Loop through each row
for (i in 1:nrow(MetaTransmit)) {
  # Check if the numeric value in Exposure_duration_numerical is contained in the string in Exposure_ID
  if (grepl(paste0("\\.", MetaTransmit$Exposure_duration_numerical[i], "\\."), MetaTransmit$Exposure_ID[i])) {
    MetaTransmit$ContainsNumeric[i] <- TRUE
  }
  # Check if ".SH." matches the value in column A and ".SH." matches the value in column C
  if (grepl("\\.SH\\.", MetaTransmit$Exposure_ID[i]) && MetaTransmit$Exposure_type[i] == "Shared") {
    MetaTransmit$MatchesExposureType[i] <- TRUE
  }
  
  # Check if ".SQ." matches the value in column A and ".SQ." matches the value in column C
  if (grepl("\\.SQ\\.", MetaTransmit$Exposure_ID[i]) && MetaTransmit$Exposure_type[i] == "Sequential") {
    MetaTransmit$MatchesExposureType[i] <- TRUE
  }
}

# Print the result
summary(MetaTransmit$MatchesExposureType)
#there were three falses, and they were all infective animals that didn't make it to the shared expt so all good. 
summary(MetaTransmit$ContainsNumeric)

#deleting the columns I used here for checks: 
MetaTransmit <- MetaTransmit %>% 
  subset(select=-c(MatchesExposureType, ContainsNumeric)) 


#checking dates are correct: 
MetaTransmitDateCheck <- MetaTransmit %>% 
  subset(Death_timing == "Euthanization") %>% 
  subset(Animal_type == "Naive")

MetaTransmitDateCheck$StartDateCorrect <- FALSE

for (x in 1:nrow(MetaTransmitDateCheck)) {
  #checking exposure start dates: 
  if ((MetaTransmitDateCheck$Exposure_type[x] == "Shared" && MetaTransmitDateCheck$Exposure_start[x] == "2021-6-27") 
      | 
      (MetaTransmitDateCheck$Exposure_type[x] == "Shared" && MetaTransmitDateCheck$Exposure_start[x] == "2021-7-11") 
      |
      (MetaTransmitDateCheck$Exposure_type[x] == "Sequential" && 
       MetaTransmitDateCheck$Exposure_duration_numerical[x] == "1" && 
       MetaTransmitDateCheck$Exposure_start[x] == "2021-7-11") 
      | 
      (MetaTransmitDateCheck$Exposure_type[x] == "Sequential" && 
       MetaTransmitDateCheck$Exposure_duration_numerical[x] == "1" && 
       MetaTransmitDateCheck$Exposure_start[x] == "2021-6-27")
      |
      (MetaTransmitDateCheck$Exposure_type[x] == "Sequential" && 
       MetaTransmitDateCheck$Exposure_duration_numerical[x] == "6" && 
       MetaTransmitDateCheck$Exposure_start[x] == "2021-7-11") 
      | 
      (MetaTransmitDateCheck$Exposure_type[x] == "Sequential" && 
       MetaTransmitDateCheck$Exposure_duration_numerical[x] == "6" && 
       MetaTransmitDateCheck$Exposure_start[x] == "2021-6-27")
      |
      (MetaTransmitDateCheck$Exposure_type[x] == "Sequential" && 
       MetaTransmitDateCheck$Exposure_duration_numerical[x] == "12" && 
       MetaTransmitDateCheck$Exposure_start[x] == "2021-7-11") 
      | 
      (MetaTransmitDateCheck$Exposure_type[x] == "Sequential" && 
       MetaTransmitDateCheck$Exposure_duration_numerical[x] == "12" && 
       MetaTransmitDateCheck$Exposure_start[x] == "2021-6-27")
      | 
      (MetaTransmitDateCheck$Exposure_type[x] == "Sequential" && 
       MetaTransmitDateCheck$Exposure_duration_numerical[x] == "24" && 
       MetaTransmitDateCheck$Exposure_start[x] == "2021-7-12") 
      | 
      (MetaTransmitDateCheck$Exposure_type[x] == "Sequential" && 
       MetaTransmitDateCheck$Exposure_duration_numerical[x] == "24" && 
       MetaTransmitDateCheck$Exposure_start[x] == "2021-6-28") 
      | 
      (MetaTransmitDateCheck$Exposure_type[x] == "Sequential" && 
       MetaTransmitDateCheck$Exposure_duration_numerical[x] == "48" && 
       MetaTransmitDateCheck$Exposure_start[x] == "2021-7-13") 
      | 
      (MetaTransmitDateCheck$Exposure_type[x] == "Sequential" && 
       MetaTransmitDateCheck$Exposure_duration_numerical[x] == "48" && 
       MetaTransmitDateCheck$Exposure_start[x] == "2021-6-29") 
      ) {
    MetaTransmitDateCheck$StartDateCorrect[x] <- TRUE
  }
}

summary(MetaTransmitDateCheck$StartDateCorrect)

#ok I found one that was recorded wrong. 
#MR438 (AHS.SQ.48.5) had a start date incorrectly written as 7/11. It should be 7/13 since this was a 48hr sequential. 
#I confirmed it was euthanized on the 18th so this was clearly an error in the entry. I corrected it in the excel doc. 

#what about death dates?

MetaTransmitDateCheck$DeathDateCorrect <- FALSE

for (y in 1:nrow(MetaTransmitDateCheck)) {
  #checking exposure start dates: 
  if( #SHARED CONDITIONS
      (MetaTransmitDateCheck$Exposure_type[y] == "Shared" && 
       MetaTransmitDateCheck$Exposure_duration_numerical[y] == "1" && 
       MetaTransmitDateCheck$Death_date[y] == "2021-07-16") 
      | 
      (MetaTransmitDateCheck$Exposure_type[y] == "Shared" && 
       MetaTransmitDateCheck$Exposure_duration_numerical[y] == "1" && 
       MetaTransmitDateCheck$Death_date[y] == "2021-07-2")
      |
      (MetaTransmitDateCheck$Exposure_type[y] == "Shared" && 
       MetaTransmitDateCheck$Exposure_duration_numerical[y] == "6" && 
       MetaTransmitDateCheck$Death_date[y] == "2021-07-16") 
      | 
      (MetaTransmitDateCheck$Exposure_type[y] == "Shared" && 
       MetaTransmitDateCheck$Exposure_duration_numerical[y] == "6" && 
       MetaTransmitDateCheck$Death_date[y] == "2021-07-02")
      |
      (MetaTransmitDateCheck$Exposure_type[y] == "Shared" && 
       MetaTransmitDateCheck$Exposure_duration_numerical[y] == "12" && 
       MetaTransmitDateCheck$Death_date[y] == "2021-07-16") 
      | 
      (MetaTransmitDateCheck$Exposure_type[y] == "Shared" && 
       MetaTransmitDateCheck$Exposure_duration_numerical[y] == "12" && 
       MetaTransmitDateCheck$Death_date[y] == "2021-07-02")
      | 
      (MetaTransmitDateCheck$Exposure_type[y] == "Shared" && 
       MetaTransmitDateCheck$Exposure_duration_numerical[y] == "24" && 
       MetaTransmitDateCheck$Death_date[y] == "2021-07-17") 
      | 
      (MetaTransmitDateCheck$Exposure_type[y] == "Shared" && 
       MetaTransmitDateCheck$Exposure_duration_numerical[y] == "24" && 
       MetaTransmitDateCheck$Death_date[y] == "2021-07-03") 
      | 
      (MetaTransmitDateCheck$Exposure_type[y] == "Shared" && 
       MetaTransmitDateCheck$Exposure_duration_numerical[y] == "48" && 
       MetaTransmitDateCheck$Death_date[y] == "2021-07-18") 
      | 
      (MetaTransmitDateCheck$Exposure_type[y] == "Shared" && 
       MetaTransmitDateCheck$Exposure_duration_numerical[y] == "48" && 
       MetaTransmitDateCheck$Death_date[y] == "2021-07-04") 
      |
      #SEQUENTIAL CONDITIONS
      
      (MetaTransmitDateCheck$Exposure_type[y] == "Sequential" && 
       MetaTransmitDateCheck$Exposure_duration_numerical[y] == "1" && 
       MetaTransmitDateCheck$Death_date[y] == "2021-07-18") 
      | 
      (MetaTransmitDateCheck$Exposure_type[y] == "Sequential" && 
       MetaTransmitDateCheck$Exposure_duration_numerical[y] == "1" && 
       MetaTransmitDateCheck$Death_date[y] == "2021-07-4")
      |
      (MetaTransmitDateCheck$Exposure_type[y] == "Sequential" && 
       MetaTransmitDateCheck$Exposure_duration_numerical[y] == "6" && 
       MetaTransmitDateCheck$Death_date[y] == "2021-07-18") 
      | 
      (MetaTransmitDateCheck$Exposure_type[y] == "Sequential" && 
       MetaTransmitDateCheck$Exposure_duration_numerical[y] == "6" && 
       MetaTransmitDateCheck$Death_date[y] == "2021-07-04")
      |
      (MetaTransmitDateCheck$Exposure_type[y] == "Sequential" && 
       MetaTransmitDateCheck$Exposure_duration_numerical[y] == "12" && 
       MetaTransmitDateCheck$Death_date[y] == "2021-07-18") 
      | 
      (MetaTransmitDateCheck$Exposure_type[y] == "Sequential" && 
       MetaTransmitDateCheck$Exposure_duration_numerical[y] == "12" && 
       MetaTransmitDateCheck$Death_date[y] == "2021-07-04")
      | 
      (MetaTransmitDateCheck$Exposure_type[y] == "Sequential" && 
       MetaTransmitDateCheck$Exposure_duration_numerical[y] == "24" && 
       MetaTransmitDateCheck$Death_date[y] == "2021-07-19") 
      | 
      (MetaTransmitDateCheck$Exposure_type[y] == "Sequential" && 
       MetaTransmitDateCheck$Exposure_duration_numerical[y] == "24" && 
       MetaTransmitDateCheck$Death_date[y] == "2021-07-05") 
      | 
      (MetaTransmitDateCheck$Exposure_type[y] == "Sequential" && 
       MetaTransmitDateCheck$Exposure_duration_numerical[y] == "48" && 
       MetaTransmitDateCheck$Death_date[y] == "2021-07-20") 
      | 
      (MetaTransmitDateCheck$Exposure_type[y] == "Sequential" && 
       MetaTransmitDateCheck$Exposure_duration_numerical[y] == "48" && 
       MetaTransmitDateCheck$Death_date[y] == "2021-07-06")
    
  ) {
    MetaTransmitDateCheck$DeathDateCorrect[y] <- TRUE
  }
}

summary(MetaTransmitDateCheck$DeathDateCorrect)
#all correct!

```


Loading in the pre-metamorph data from Nicole. 
This is the masses and metamorphosis dates. 

```{r}

library(dplyr)
pre_exposure <- read_csv("form-1__pre-exposure-animals.csv")

pre_exposure <- pre_exposure %>% 
  rename(Container_ID = `1_Container_ID_ex_EL`, 
         Animal_type = `2_Was_the_animal_a_v`,
         Larval_tank = `3_Larval_Tank`,
         Date_metamorphed = `4_Date_metamorphed`, 
         Date_tail_absorbed = `5_Date_tail_absorbed`, 
         Mass_at_mass_date = `6_Mass_at_mass_date_`, 
         SVL_at_mass_date = `7_SVL_at_mass_date`, 
         Date_group_housed = `8_Date_communally_ho`) %>% 
  select(-c(ec5_uuid, uploaded_at, title, created_by)) %>% 
  filter(!Animal_type == "Virus") %>% 
  mutate(Date_metamorphed = dmy(Date_metamorphed), 
         Date_tail_absorbed = dmy(Date_tail_absorbed), 
         Date_group_housed = dmy(Date_group_housed))

class(pre_exposure$Date_metamorphed)

#were there differences between the masses of the two batches? 

ggplot(pre_exposure, aes(x=Date_tail_absorbed, y=Mass_at_mass_date)) +
  geom_point() +
  geom_smooth(method = "lm", color = "blue")

#there are a couple of outlier points that definitely don't make sense. 
#check to see if the metamorph date is a little bit before tail absorption, and if group housed date is after tail absorption
#so: Date_metamorphed < Date_tail_absorbed < Date_group_housed

pre_exposure <- pre_exposure %>% 
    mutate(Dates_in_order = Date_metamorphed <= Date_tail_absorbed & 
             Date_tail_absorbed <= Date_group_housed)

summary(pre_exposure$Dates_in_order)

wrong_dates <- pre_exposure %>% 
  filter(Dates_in_order == FALSE)

#3 dates are out of order. I think that I will just use date tail absorbed instead of date metamorphed

#SVL
ggplot(pre_exposure, aes(x=Date_tail_absorbed, y=SVL_at_mass_date)) +
  geom_point() +
  geom_smooth(method = "lm", color = "blue")

#what are the batch start dates for my experiment? 
summary(MetaTransmit$Exposure_start[MetaTransmit$Batch == 1])
#batch 1 starts 6/27 - 6/29

summary(MetaTransmit$Exposure_start[MetaTransmit$Batch == 2])
#batch 2 starts 7/11 - 7/13

#can I figure out exactly which batch was which using the pre-exp IDs / container IDs? 
MetaTransmit$Pre_exp_ID <- as.factor(MetaTransmit$Pre_exp_ID)
summary(MetaTransmit$Pre_exp_ID)

pre_exposure$Container_ID <- as.factor(pre_exposure$Container_ID)
summary(pre_exposure$Container_ID)

Container_ID_summaries <- pre_exposure %>% 
  select(Date_metamorphed, Container_ID, Date_tail_absorbed, Date_group_housed, 
         Mass_at_mass_date, SVL_at_mass_date) %>% 
  dplyr::group_by(Container_ID) %>% 
  dplyr::summarize(
    # Date range for Date_metamorphed
    date_meta_start = min(Date_metamorphed, na.rm = TRUE),
    date_meta_mean = mean(Date_metamorphed, na.rm = TRUE),
    date_meta_end = max(Date_metamorphed, na.rm = TRUE),
    # Date range for Date_tail_absorbed
    date_tail_absorbed_start = min(Date_tail_absorbed, na.rm = TRUE),
    date_tail_absorbed_mean = mean(Date_tail_absorbed, na.rm = TRUE),
    date_tail_absorbed_end = max(Date_tail_absorbed, na.rm = TRUE),
    # Date range for Date_metamorphed
    date_housed_start = min(Date_group_housed, na.rm = TRUE),
    date_housed_mean = mean(Date_group_housed, na.rm = TRUE),
    date_housed_end = max(Date_group_housed, na.rm = TRUE),
    # Additional summary statistics
    avg_mass = mean(Mass_at_mass_date, na.rm = TRUE),
    avg_svl = mean(SVL_at_mass_date, na.rm = TRUE),
    count = n()
  )

#now lets see if we can get a list of pre exposure IDs from meta transmit by batch
MetaTransmit$Batch <- as.numeric(MetaTransmit$Batch)

Pre_exp_ID_summaries <- MetaTransmit %>% 
  select(Pre_exp_ID, Batch) %>% 
  group_by(Pre_exp_ID) %>% 
  dplyr::summarize(
    Batch_consensus = mean(Batch, na.rm = TRUE), 
    count = n()
  )

Pre_exp_ID_summaries <- Pre_exp_ID_summaries %>% 
  mutate(Batch_known = case_when(
    Batch_consensus < 2.01 & Batch_consensus > 1.99 ~ 2, 
    Batch_consensus < 1.01 & Batch_consensus > 0.99 ~ 1, 
    .default = NA
  ), 
  Batch_approx = round(Batch_consensus, 0))

Joined_ID_summaries <- Container_ID_summaries %>% 
  left_join(Pre_exp_ID_summaries, by = join_by(Container_ID == Pre_exp_ID))


#is there a difference in mass and SVL by batch? 
#MASS
Joined_ID_summaries$Batch_known <- as.factor(Joined_ID_summaries$Batch_known)

ggplot(Joined_ID_summaries, aes(y=avg_mass, x=Batch_known, fill = Batch_known)) +
  geom_boxplot() +
  scale_fill_brewer(palette = "Set3")

t_test_mass <- t.test(Joined_ID_summaries$avg_mass[Joined_ID_summaries$Batch_known == 1], 
                      Joined_ID_summaries$avg_mass[Joined_ID_summaries$Batch_known == 2], 
                      var.equal = TRUE)
print(t_test_mass)

#the differences in means are statistically significant. Reject null hypothesis. p-value = 0.004223

#but let's also check by weighting the T tests by our counts. 
sum(Joined_ID_summaries$count.y, na.rm = TRUE)
sum(Joined_ID_summaries$count.x, na.rm = TRUE)

library(weights)
weighted_t_test_mass <- wtd.t.test(
  x = Joined_ID_summaries$avg_mass[Joined_ID_summaries$Batch_known == 1],
  y = Joined_ID_summaries$avg_mass[Joined_ID_summaries$Batch_known == 2],
  weight = Joined_ID_summaries$count.y[Joined_ID_summaries$Batch_known == 1],
  weighty = Joined_ID_summaries$count.y[Joined_ID_summaries$Batch_known == 2]
)

print(weighted_t_test_mass)

#also significant. p = 0.001509168. Reject the null, so the groups are different. 
#batch 1 = 0.29482312 g
#batch 2 = 0.34797262 g

#SVL
ggplot(Joined_ID_summaries, aes(y=avg_svl, x=Batch_known, fill = Batch_known)) +
  geom_boxplot() +
  scale_fill_brewer(palette = "Set3")

t_test_svl <- t.test(Joined_ID_summaries$avg_svl[Joined_ID_summaries$Batch_known == 1], 
                      Joined_ID_summaries$avg_svl[Joined_ID_summaries$Batch_known == 2], 
                      var.equal = TRUE)
print(t_test_svl)

#yep SVL is significantly different. reject the null. 
#batch 1 SVL mean = 13.74603
#batch 2 SVL mean = 14.26265

#weighted t test
weighted_t_test_svl <- wtd.t.test(
  x = Joined_ID_summaries$avg_svl[Joined_ID_summaries$Batch_known == 1],
  y = Joined_ID_summaries$avg_svl[Joined_ID_summaries$Batch_known == 2],
  weight = Joined_ID_summaries$count.y[Joined_ID_summaries$Batch_known == 1],
  weighty = Joined_ID_summaries$count.y[Joined_ID_summaries$Batch_known == 2]
)

print(weighted_t_test_svl)

#significant p value. 0.001448863. reject the null. 
#batch 1 SVL mean = 13.619
#batch 2 SVL mean = 14.226

MetaTransmit <- MetaTransmit %>% 
  left_join(Joined_ID_summaries, by = join_by(Pre_exp_ID == Container_ID))
```






Subseting out a few smaller data sets for future analysis. 
Specifically: 
 - Naive
     - Naive Sequential
       - Zeroes removed for viral load
     - Naive Shared
       - Zeroes removed for viral load
 - Infected
     - zeroes removed for viral load
 - Control animals
     - Control pseudo-naive
     - Control pseudo-infected 
 
```{r}
#----------Subset naive data
MetaTransmit_naive <- MetaTransmit %>% 
  subset(Animal_type== "Naive",)

#Subset Sequential Naives
MetaTransmitNaiveSQ <- MetaTransmit_naive %>% 
  subset(Exposure_type == "Sequential",)
#subset Shared Naives
MetaTransmitNaiveSH <- MetaTransmit_naive %>% 
  subset(Exposure_type == "Shared",)

# --------Getting rid of 0s for the shared and sequential data

#Naives Overall
NoZeroesNaive <- MetaTransmit %>%
  subset(Animal_type == "Naive")%>%
  subset(!(MeanCopyFinal == 0)) %>%
  mutate(logMeanCopyFinal = log(MeanCopyFinal))

#SHARED
#getting rid of the 0s as recommended
NoZeroesNaiveSH <- MetaTransmitNaiveSH %>% 
  subset(!(MeanCopyFinal == 0)) %>% 
  #Adding a variable called logmeancopyfinal without the zeroes
  mutate(logMeanCopyFinal = log(MeanCopyFinal)) 

#SEQUENTIAL
#getting rid of the 0s as Jesse Recommended
NoZeroesNaiveSQ <- MetaTransmitNaiveSQ %>% 
  subset(!(MeanCopyFinal == 0)) %>% 
  #Adding a variable called logmeancopyfinal without the zeroes
  mutate(logMeanCopyFinal = log(MeanCopyFinal)) 

#-----------Subseting out virus animals
MetaTransmit_virus <- MetaTransmit %>% 
  subset(MetaTransmit$Animal_type=="Infected",)

#getting rid of the 0s for viral load analysis
NoZeroesVirus <- MetaTransmit_virus %>% 
  subset(!(MeanCopyFinal == 0)) %>% 
  #Adding a variable called logmeancopyfinal without the zeroes
  mutate(logMeanCopyFinal = log(MeanCopyFinal)) 



#------------Subsetting  control animals 
MetaTransmit_controlPN <- MetaTransmit %>% 
  subset(Animal_type=="Control (pseudo-naive)",)
MetaTransmit_controlPV <- MetaTransmit %>% 
  subset(MetaTransmit$Animal_type=="Control (pseudo-infected)",)
```

## The Plots

Naive Suspected Mortalities Based on Death Timing

```{r}
#Suspected Ranavirus Mortalities (Focal Naive)

#facet labels
# New facet label names for exposure type variable
exposure.labs <- c("Sequential habitation", "Cohabitation")
names(exposure.labs) <- c("Sequential", "Shared")

MetaTransmit_naive$Death_timing <- as.factor(MetaTransmit_naive$Death_timing)
MetaTransmit_naive$Exposure_duration_numerical <- as.factor(MetaTransmit_naive$Exposure_duration_numerical)
MetaTransmit_naive$Exposure_type <- as.factor(MetaTransmit_naive$Exposure_type)
MetaTransmit_naive$Exposure_duration_numerical <- as.factor(MetaTransmit_naive$Exposure_duration_numerical)

#Fixed the graph to show proportions instead of counts: 

#annotations for the text on the ggplot
annotations <- data.frame(x1 = c(1, 1, 1, 4, 4, 4), 
                   y1 = c(.85, 0.69, 0.4, .95, .86, .5), 
                   Death_timing = c("Died after","Died during", "Euthanized", 
                                    "Died after","Died during", "Euthanized"),
                   Exposure_type = c("Sequential", "Sequential", "Sequential", 
                                     "Shared", "Shared", "Shared")
                   )

ggplot(data = MetaTransmit_naive, aes(x = Exposure_duration_numerical, fill = Death_timing)) + 
  geom_bar(position = "fill") +  # This will normalize the heights of the bars to show proportions
  guides(fill = guide_legend(title = NULL)) + 
  facet_wrap(.~Exposure_type, labeller = labeller(Exposure_type = exposure.labs)) +
  geom_text(data = annotations, aes(x = x1, y = y1, label = Death_timing), 
            size = 2.8, family = "helvetica", 
            color = c("white", "black", "white", "white", "black", "white"), 
            fontface = "bold") +
  labs(x = "Time between exposures (hrs)               Exposure period (hrs)", 
       y = "Proportion") +  # Changed y label to reflect proportions
  theme(text = element_text(size = 2),
        axis.title.y = element_text(size = 15, hjust = 0.7, family = "helvetica"),
        axis.title.x = element_text(size = 15, hjust = 0.3, family = "helvetica"),
        strip.text.x = element_text(size = 20, family = "helvetica"),
        axis.text.x = element_text(size = 15, family = "helvetica"),
        axis.text.y = element_text(size = 15, family = "helvetica"),
        legend.position = "none") + 
  scale_fill_manual(values=c("#A85F82", "#FEFFF3","#000025", "#A85F82", "#FEFFF3","#000025"))
  
ggsave(
  'suspectedMortalities.jpg', plot = last_plot(), width = 200, height = 120, units = "mm")

```


Reviewer requested a plot of premature deaths for infective animals too. 

Infective Suspected Mortalities Based on Death Timing

```{r}
#Suspected Ranavirus Mortalities (Infectives)

#facet labels
# New facet label names for exposure type variable
exposure.labs <- c("Sequential habitation", "Cohabitation")
names(exposure.labs) <- c("Sequential", "Shared")

MetaTransmit_virus$Death_timing <- as.factor(MetaTransmit_virus$Death_timing)
MetaTransmit_virus$Exposure_duration_numerical <- as.factor(MetaTransmit_virus$Exposure_duration_numerical)
MetaTransmit_virus$Exposure_type <- as.factor(MetaTransmit_virus$Exposure_type)
MetaTransmit_virus$Exposure_duration_numerical <- as.factor(MetaTransmit_virus$Exposure_duration_numerical)

#Fixed the graph to show proportions instead of counts: 

#annotations for the text on the ggplot
annotations <- data.frame(x1 = c(4, 4), 
                   y1 = c(.9, .4), 
                   Death_timing = c("Died during", "Euthanized")
                   )

ggplot(data = MetaTransmit_virus, aes(x = Exposure_duration_numerical, fill = Death_timing)) + 
  geom_bar(position = "fill") +  # This will normalize the heights of the bars to show proportions
  guides(fill = guide_legend(title = NULL)) + 
  #facet_wrap(.~Exposure_type, labeller = labeller(Exposure_type = exposure.labs)) +
  geom_text(data = annotations, aes(x = x1, y = y1, label = Death_timing), 
            size = 2.8, family = "helvetica", 
            color = c("black", "white"), 
            fontface = "bold") +
  labs(x = "Exposure period (hrs)", 
       y = "Proportion") +  # Changed y label to reflect proportions
  theme(text = element_text(size = 2),
        axis.title.y = element_text(size = 15, hjust = 0.5, family = "helvetica"),
        axis.title.x = element_text(size = 15, hjust = 0.5, family = "helvetica"),
        strip.text.x = element_text(size = 20, family = "helvetica"),
        axis.text.x = element_text(size = 15, family = "helvetica"),
        axis.text.y = element_text(size = 15, family = "helvetica"),
        legend.position = "none") + 
  scale_fill_manual(values=c("#FEFFF3","#000025", "#FEFFF3","#000025"))
  
ggsave(
  'InfectiveSuspectedMortalities.jpg', plot = last_plot(), width = 200, height = 120, units = "mm")

#summary stats for this? 
summary(MetaTransmit_virus$Death_timing)
12/127


```







Naive Ranavirus Status
This is much higher than I had expected. Nearly all of the animals got infected. I would have expected the sequential exposures to have more negatives for the higher durations (48hrs for example), instead of the lower values (1hr). 
```{r}
#Infection Status (Focal Naive)

#annotations
annotations <- data.frame(x1 = c(1, 1, 1, 1, 1, 3), 
                   y1 = c(.95, .5, .02, .98, .5, .02), 
                   Final_status = c("Neg","Pos", "NA", "Neg","Pos", "NA"),
                   Exposure_type = c("Sequential", "Sequential", "Sequential", "Shared", "Shared", "Shared")
                   )

#plot naive animal mortality
#ggplot(MetaTransmit_naive, aes(x= Exposure_duration_numerical, y= lengths(Exposure_duration_numerical), fill = Final_status)) + 
ggplot(MetaTransmit_naive, aes(x= Exposure_duration_numerical, fill = Final_status)) + 
  #stat_count(aes(x = Exposure_duration_numerical, group = Final_status)) +
  geom_bar(position = "fill") +
  guides(fill=guide_legend(title=NULL)) + 
  facet_wrap(~Exposure_type, 
  labeller = labeller(Exposure_type = exposure.labs)
  ) +
  geom_text(data = annotations, aes(x = x1,  y = y1, label = Final_status), size = 3, family = "helvetica", 
            color = c("black", "white", "white", "black", "white", "white"), fontface = "bold") +
  labs(x = "Time between exposures (hrs)               Exposure period (hrs)", y = "Susceptible animal proportion") + 
  theme(axis.title.y = element_text(size = 15, hjust = 0.7, family = "helvetica"),
        axis.title.x = element_text(size = 15, hjust = 0.3, family = "helvetica"),
        strip.text.x = element_text(size = 20, family = "helvetica"),
        axis.text.x = element_text(size = 15, family = "helvetica"),
        axis.text.y = element_text(size = 15, family = "helvetica"),
        legend.text=element_text(size=13, family = "helvetica"),
        legend.position = "none") + 
  scale_fill_manual(breaks = c("Neg", "Pos", NA), 
                       values=c("lightblue", "darkred", "grey"))

ggsave(
  'virusStatusNaive.jpg',
  plot = last_plot(), width = 200, height = 120, units = "mm"
)

```
Virus Animals
These results look good for the most part, but that one 1hr shared negative should not be negative. I think it's likely a lab error. 
```{r}
#plot mortality
ggplot(MetaTransmit_virus, aes(x= Exposure_duration , y= lengths(Exposure_duration), fill = Final_status)) + 
  geom_bar(stat= "identity") +
  guides(fill=guide_legend(title=NULL)) + 
  labs(x = "Exposure Period", y = "Infective Animal Counts") + 
  theme(axis.title.y = element_text(size = 18, hjust = 0.7),
        axis.title.x = element_text(size = 18, hjust = 0.5),
        strip.text.x = element_text(size = 20),
        axis.text.x = element_text(size = 15),
        axis.text.y = element_text(size = 15),
        legend.text=element_text(size=15),
        axis.title = element_text(size=20, color = "black"),
        legend.position = "bottom") + 
  #scale_fill_brewer(palette="Set2")
  scale_fill_manual(breaks = c("Neg", "Pos", NA), 
                       values=c("lightblue", "darkred", "grey"))

ggsave(
  'virusStatusVirus.jpg',
  plot = last_plot(),
  device = NULL,
  path = NULL,
  scale = 1,
  width = NA,
  height = NA,
  units = c("in", "cm", "mm", "px"),
  dpi = 300,
  limitsize = TRUE,
  bg = NULL
)

```


Control Naive Animals
These results are exactly as expected. None of the control animals were infected. 

Just realized 1/11/2023 that Control animal MR114, AHS.SQ.48.C pseudo-naive was incorrectly labeled as a Naive animal. I corrected it to reclassify from Naive to Control (pseudo-naive). Unfortunately, this animal did test positive, so I'll need to correct the control write up in the paper. Low viral copy number of about 100 though, and neither pair tested positive. Batch 2. I found the tape label picture and I didn't have any notes on it so I don't think it got paired with a non-control animal during the experiment, but I do have a vague memory of a control being used in an exposure experiment, but I may be thinking of something else (possibly that an ELS animal was paired with an EHS animal in the controls, which I did see in the notes). 

```{r}
#plot naive animal mortality
ggplot(MetaTransmit_controlPN, aes(x= Exposure_duration , y= lengths(Exposure_duration), fill = Final_status)) + 
  geom_bar(stat= "identity") +
  guides(fill=guide_legend(title=NULL)) + 
  facet_wrap(~Exposure_type, 
  labeller = labeller(Exposure_type = exposure.labs)
  ) +
  labs(x = "Time Between Exposures         Exposure Period", y = "Control Susceptible Counts") + 
  theme(axis.title.y = element_text(size = 18, hjust = 0.7),
        axis.title.x = element_text(size = 18, hjust = 0.3),
        strip.text.x = element_text(size = 20),
        axis.text.x = element_text(size = 15),
        axis.text.y = element_text(size = 15),
        legend.text=element_text(size=15),
        axis.title = element_text(size=20, color = "black"),
        legend.position = "bottom") + 
  #scale_fill_brewer(palette="Set2")
  scale_fill_manual(breaks = c("Neg", "Pos", NA), 
                       values=c("lightblue", "darkred", "grey"))

ggsave(
  'virusStatusControlPseudonaive.jpg',
  plot = last_plot(),
  device = NULL,
  path = NULL,
  scale = 1,
  width = NA,
  height = NA,
  units = c("in", "cm", "mm", "px"),
  dpi = 300,
  limitsize = TRUE,
  bg = NULL
)

```

Control Pseudovirus Animals: 
Note - there is one 48hr control pseudo-infected positive but the viral copy is quite low (mean = 772), so it seems like a contamination issue in the lab. It may be worth redoing the sample if that is a possibility. 
```{r}
#plot naive animal mortality
ggplot(MetaTransmit_controlPV, aes(x= Exposure_duration , y= lengths(Exposure_duration), fill = Final_status)) + 
  geom_bar(stat= "identity") +
  guides(fill=guide_legend(title=NULL)) + 
  labs(x = "Exposure Period", y = "Control Pseudo-infectives Counts") + 
  theme(axis.title.y = element_text(size = 15, hjust = 0.7),
        axis.title.x = element_text(size = 18, hjust = 0.5),
        strip.text.x = element_text(size = 20),
        axis.text.x = element_text(size = 15),
        axis.text.y = element_text(size = 15),
        legend.text=element_text(size=15),
        axis.title = element_text(size=20, color = "black"),
        legend.position = "bottom") + 
  #scale_fill_brewer(palette="Set2")
  scale_fill_manual(breaks = c("Neg", "Pos", NA), 
                       values=c("lightblue", "darkred", "grey"))

ggsave(
  'virusStatusControlPseudovirus.jpg',
  plot = last_plot(),
  device = NULL,
  path = NULL,
  scale = 1,
  width = NA,
  height = NA,
  units = c("in", "cm", "mm", "px"),
  dpi = 300,
  limitsize = TRUE,
  bg = NULL
)
```

#Performing a few data visualizations on viral load and exposure duration: 

Performing data visualization for shared and sequential boxplots and the facet-wrapped violin plots using the log mean copy final and 0s excluded (previously it was log mean copy final plus 1). Since the linear model does it this way, I think it's most appropriate to exclude 0s on the visualization. 

```{r}
#-------------Boxplots

#Violin Box Plot Shared
ggplot(NoZeroesNaiveSH, aes(x=Exposure_duration, y= log(MeanCopyFinal)))+
  geom_violin()+
  geom_boxplot(alpha=0.3) + 
  labs(x = "Exposure Period", y = "Log of Viral Load") 
ggsave("NaiveSHViralLoadViolinNoZeroes.jpg", plot = last_plot())

#Violin Box Plot Sequential
ggplot(NoZeroesNaiveSQ, aes(x=Exposure_duration, y= log(MeanCopyFinal)))+
  geom_violin()+
  geom_boxplot(alpha=0.3) + 
  labs(x = "Time Between Exposures", y = "Log of Viral Load")
ggsave("NaiveSQViralLoadViolinNoZeroes.jpg", plot = last_plot())

NoZeroesNaive$Exposure_duration_numerical <- as.factor(NoZeroesNaive$Exposure_duration_numerical)

#violin boxplots of shared and sequential facet_wrapped
ggplot(NoZeroesNaive, aes(x=Exposure_duration_numerical, y= MeanCopyFinal, 
                               fill = Exposure_duration_numerical)) +
  #geom_violin()+
  geom_boxplot(alpha=0.5, outlier.shape = NA) + 
  geom_jitter(size = 0.8, alpha = 0.5, position = position_jitter(width = .2)) +
  scale_y_continuous(trans='log10') +
  annotation_logticks(sides = "l", outside = TRUE, 
                      short = unit(.4,"mm"),
                      mid = unit(1,"mm"),
                      long = unit(2,"mm")) +
  coord_cartesian(clip = "off") +
  facet_wrap(~Exposure_type, 
  labeller = labeller(Exposure_type = exposure.labs)
  ) +
  labs(x = "Time between exposures (hrs)               Exposure period (hrs)", y = "Viral load") + 
  theme(axis.title.y = element_text(size = 15, hjust = 0.7, family = "helvetica"),
        axis.title.x = element_text(size = 15, hjust = 0.3, family = "helvetica"),
        strip.text.x = element_text(size = 20, family = "helvetica"),
        axis.text.x = element_text(size = 15, family = "helvetica"),
        axis.text.y = element_text(size = 15, family = "helvetica"),
        #axis.ticks.length.y = unit(4, "mm"),
        legend.position = "none") + 
  scale_fill_brewer(palette = "YlOrRd")
ggsave("ViralLoadBoxNoZeroes.jpg", plot = last_plot(), width = 200, height = 120, units = "mm")
  
```



Reviewer requested estimates of the infective frog viral loads. Using the same code as above, just with the infectives. 

```{r}

NoZeroesVirus$Exposure_duration_numerical <- as.factor(NoZeroesVirus$Exposure_duration_numerical)

#boxplots of shared and sequential facet_wrapped
ggplot(NoZeroesVirus, aes(x=Exposure_duration_numerical, y= MeanCopyFinal, 
                               fill = Exposure_duration_numerical)) +
  #geom_violin()+
  geom_boxplot(alpha=0.5, outlier.shape = NA) + 
  geom_jitter(size = 0.8, alpha = 0.5, position = position_jitter(width = .2)) +
  scale_y_continuous(trans='log10') +
  annotation_logticks(sides = "l", outside = TRUE, 
                      short = unit(.4,"mm"),
                      mid = unit(1,"mm"),
                      long = unit(2,"mm")) +
  coord_cartesian(clip = "off") +
  #facet_wrap(~Exposure_type, 
  #labeller = labeller(Exposure_type = exposure.labs)
  #) +
  labs(x = "Exposure period (hrs)", y = "Viral load") + 
  theme(axis.title.y = element_text(size = 15, hjust = 0.5, family = "helvetica"),
        axis.title.x = element_text(size = 15, hjust = 0.5, family = "helvetica"),
        strip.text.x = element_text(size = 20, family = "helvetica"),
        axis.text.x = element_text(size = 15, family = "helvetica"),
        axis.text.y = element_text(size = 15, family = "helvetica"),
        #axis.ticks.length.y = unit(4, "mm"),
        legend.position = "none") + 
  scale_fill_brewer(palette = "YlOrRd")
ggsave("VirusViralLoadBoxNoZeroes.jpg", plot = last_plot(), width = 200, height = 120, units = "mm")
```


#Linear Models for Viral Load and Exposure Durations by Experiment: 

```{r}

#--------------NEW ATTEMPT BASED ON JESSE BRUNNER'S FEEDBACK 4/20
#assuming no intercept
#V ≈ b*Duration
#log(V) ≈ log(b) + log(Duration)

NoZeroesNaiveSH$Exposure_duration_numerical <- as.numeric(NoZeroesNaiveSH$Exposure_duration_numerical)
NoZeroesNaiveSQ$Exposure_duration_numerical <- as.numeric(NoZeroesNaiveSQ$Exposure_duration_numerical)


#the linear model shared
log_transformed_lm_SH = glm(log(MeanCopyFinal) ~ log(Exposure_duration_numerical), data = NoZeroesNaiveSH)
summary(log_transformed_lm_SH)
plot(log_transformed_lm_SH)

#the linear model sequential
log_transformed_lm_SQ = glm(log(MeanCopyFinal) ~ log(Exposure_duration_numerical), data = NoZeroesNaiveSQ)
summary(log_transformed_lm_SQ)
plot(log_transformed_lm_SQ)
```


#Paired Pearson's Linear Model: 

The goal here is to compare the viral load of the infected animal with the paired naive animal to see if an animal gets a more severe infection from an infected animal with a higher viral load. 

```{r}
#Ok so this is pairing the infected animal with the naive animal by ID for the Shared housing experiment. 
class(MetaTransmit$avg_mass)

#Shared Housing
Shared_grouped <- MetaTransmit %>% 
  select(Trio_ID, Exposure_type, MeanCopyFinal, Animal_type, Exposure_duration_numerical, Larval_condition, Batch) %>% 
  filter((Exposure_type == "Shared" & Animal_type == "Infected") | 
           (Exposure_type == "Shared" & Animal_type == "Naive")) %>% 
  select(Trio_ID, Animal_type, MeanCopyFinal, Exposure_duration_numerical, Larval_condition, Batch) %>% 
  pivot_wider(names_from = Animal_type, values_from = MeanCopyFinal)

# Separate avg_mass for "Naive" before pivoting
metamorph_stats_naive_SH <- MetaTransmit %>%
  filter(Exposure_type == "Shared" & Animal_type == "Naive") %>%
  select(Trio_ID, avg_mass, avg_svl)

#add it back in
Shared_grouped <- Shared_grouped %>% 
  left_join(metamorph_stats_naive_SH, by = "Trio_ID")

#Sequential Housing
Sequential_grouped <- MetaTransmit %>% 
  select(Trio_ID, Exposure_type, MeanCopyFinal, Animal_type, Exposure_duration_numerical, Larval_condition, Batch) %>% 
  filter((Exposure_type == "Shared" & Animal_type == "Infected") | 
           (Exposure_type == "Sequential" & Animal_type == "Naive")) %>% 
  select(Trio_ID, Animal_type, MeanCopyFinal, Exposure_duration_numerical, Larval_condition, Batch) %>% 
  pivot_wider(names_from = Animal_type, values_from = MeanCopyFinal)

# Separate avg_mass for "Naive" before pivoting
metamorph_stats_naive_SQ <- MetaTransmit %>%
  filter(Exposure_type == "Sequential" & Animal_type == "Naive") %>%
  select(Trio_ID, avg_mass, avg_svl)

#add it back in
Sequential_grouped <- Sequential_grouped %>% 
  left_join(metamorph_stats_naive_SQ, by = "Trio_ID")

#---------------------------------------
#Can I group the data better?
merged_groups <- merge(Shared_grouped, Sequential_grouped, all.x = TRUE, all.y = TRUE)

Grouped_data <- MetaTransmit %>% 
  select(Trio_ID, Exposure_type, MeanCopyFinal, Animal_type, Exposure_duration_numerical, Larval_condition, Batch) %>% 
  filter(Exposure_type == "Shared" & Animal_type == "Infected" | 
           Exposure_type == "Shared" & Animal_type == "Naive" | 
           Exposure_type == "Sequential" & Animal_type == "Naive") %>% 
  select(Trio_ID, Animal_type, Exposure_type, MeanCopyFinal, Exposure_duration_numerical, Larval_condition, Batch) %>% 
  pivot_wider(names_from = Animal_type | Exposure_type, values_from = MeanCopyFinal)

#doesn't have larval condition or batch
#---------------------------------------------------------
#trying to run GLMs on the paired data: need to subset out the 0s
Shared_grouped_nozeroes <- Shared_grouped %>% 
  subset(!(Naive == 0 | Infected == 0)) #| is.na(Naive) | is.na(Infected)))

Sequential_grouped_nozeroes <- Sequential_grouped %>% 
  subset(!(Naive == 0 | Infected == 0))


#SHARED
Shared_grouped_nozeroes$Exposure_duration_numerical <- as.numeric(Shared_grouped_nozeroes$Exposure_duration_numerical)

#all the vars
lm_SH_all = glm(log(Naive) ~ log(Exposure_duration_numerical) + log(Infected) + Batch + Larval_condition + avg_mass + avg_svl, data = Shared_grouped_nozeroes)
summary(lm_SH_all)
plot(lm_SH_all)


#all 4 variables
lm_SH_all_4 = glm(log(Naive) ~ log(Exposure_duration_numerical) + log(Infected) + Batch + Larval_condition, data = Shared_grouped_nozeroes)
summary(lm_SH_all_4)
plot(lm_SH_all_4)
#AIC = 674.24
#here, larval_condition is not significant, and neither is log_infected. 

#take out larval condition since it isn't significant. Does it improve AIC?
lm_SH_no_larval = glm(log(Naive) ~ log(Exposure_duration_numerical) + log(Infected) + Batch, data = Shared_grouped_nozeroes)
summary(lm_SH_no_larval)
plot(lm_SH_no_larval)
#yes, AIC improved. AIC = 671.32

#separately, take out log(infected) since it isn't significant. Does it improve AIC compared to original?
lm_SH_no_infected = glm(log(Naive) ~ log(Exposure_duration_numerical) + Batch + Larval_condition, data = Shared_grouped_nozeroes)
summary(lm_SH_no_infected)
plot(lm_SH_no_infected)
#yes, AIC improved. AIC = 672.6

#now take out both log(infected) and larval condition, since these separately improved the model. 
lm_SH_exposure_and_batch = glm(log(Naive) ~ log(Exposure_duration_numerical) + Batch, data = Shared_grouped_nozeroes)
summary(lm_SH_exposure_and_batch)
plot(lm_SH_exposure_and_batch)
#yes, together these improved the AIC. AIC = 669.43
#BEST MODEL

#testing just exposure. (predicted to be worse. this is verification)
lm_SH_exposure = glm(log(Naive) ~ log(Exposure_duration_numerical) , data = Shared_grouped_nozeroes)
summary(lm_SH_exposure)
plot(lm_SH_exposure)
#worse model. AIC = 692.61

#testing only batch (predicted to be worse. this is verification)
lm_SH_batch = glm(log(Naive) ~ Batch, data = Shared_grouped_nozeroes)
summary(lm_SH_batch)
plot(lm_SH_batch)
#worse model. AIC = 704.64


#individual variable tests
#also removing the 2 significant vars with an n-1 approach
lm_SH_no_batch = glm(log(Naive) ~ log(Exposure_duration_numerical) + log(Infected) + Larval_condition, data = Shared_grouped_nozeroes)
summary(lm_SH_no_batch)
#predicted correctly to be worse. AIC = 697.19

#also removing the 2 significant vars with an n-1 approach
lm_SH_no_exposure = glm(log(Naive) ~ log(Infected) + Batch + Larval_condition, data = Shared_grouped_nozeroes)
summary(lm_SH_no_exposure)
#also worse. AIC= 696.74


#3 variable model <-- USING THIS ONE
lm_SH_all_3 = glm(log(Naive) ~ log(Exposure_duration_numerical) + log(Infected) + Batch, data = Shared_grouped_nozeroes)
summary(lm_SH_all_3)
plot(lm_SH_all_3)

#5 var model (adding in mass and svl at metamorphosis)
lm_SH_5 = glm(log(Naive) ~ log(Exposure_duration_numerical) + log(Infected) + Batch + avg_mass + avg_svl, data = Shared_grouped_nozeroes)
summary(lm_SH_5)
plot(lm_SH_5)

#ok larval mass at metamorphosis and SVL are not significant predictors, even though batch still is. 
#this means that even though the batch 2 frogs are larger at metamorphosis, that is not the reason for the higher viral loads in Batch 2. 


#-----------------------------
#SEQUENTIAL
Sequential_grouped_nozeroes$Exposure_duration_numerical <- as.numeric(Sequential_grouped_nozeroes$Exposure_duration_numerical)

#all 4 variables
lm_SQ_all_4 = glm(log(Naive) ~ log(Exposure_duration_numerical) + log(Infected) + Batch + Larval_condition, data = Sequential_grouped_nozeroes)
summary(lm_SQ_all_4)
#AIC: 676.36
#batch and larval condition not significant. log(infected) is borderline

#take out batch --> improved model AIC
lm_SQ_no_batch = glm(log(Naive) ~ log(Exposure_duration_numerical) + log(Infected) + Larval_condition, data = Sequential_grouped_nozeroes)
summary(lm_SQ_no_batch)
#AIC: 674.65

#separately, take out larval condition --> improved model AIC
lm_SQ_no_larval = glm(log(Naive) ~ log(Exposure_duration_numerical) + log(Infected) + Batch, data = Sequential_grouped_nozeroes)
summary(lm_SQ_no_larval)
#AIC: 673.71

#combine the 2 - no larval and no batch. It's slightly better. 
lm_SQ_exposure_and_infected = glm(log(Naive) ~ log(Exposure_duration_numerical) + log(Infected), data = Sequential_grouped_nozeroes)
summary(lm_SQ_exposure_and_infected)
#AIC: 672.15
#BEST MODEL

#testing eliminating the significant and borderline sig variables (combinations of 3): should be worse
#no log infected --> worse model AIC
lm_SQ_no_infected = glm(log(Naive) ~ log(Exposure_duration_numerical) + Batch + Larval_condition, data = Sequential_grouped_nozeroes)
summary(lm_SQ_no_infected)
#AIC: 677.84

#no exposure duration  --> worse model AIC
lm_SQ_no_exposure = glm(log(Naive) ~ log(Infected) + Batch + Larval_condition, data = Sequential_grouped_nozeroes)
summary(lm_SQ_no_exposure)
#AIC: 679.85


#what about individual variables (test just one?)
lm_SQ_exposure = glm(log(Naive) ~ log(Exposure_duration_numerical), data = Sequential_grouped_nozeroes)
summary(lm_SQ_exposure)
#AIC: 672.39
#it's really really close, but including infected does improve the model slightly

lm_SQ_infected = glm(log(Naive) ~ log(Infected), data = Sequential_grouped_nozeroes)
summary(lm_SQ_infected)
#AIC: 676.59
#this is worse 

#testing metamorphosis mass and svl
#5 var model (adding in mass and svl at metamorphosis to the 3 var model)
lm_SQ_5 = glm(log(Naive) ~ log(Exposure_duration_numerical) + log(Infected) + Batch + avg_mass + avg_svl, data = Sequential_grouped_nozeroes)
summary(lm_SQ_5)
plot(lm_SQ_5)

#ok larval mass at metamorphosis and SVL are not significant predictors, even though batch still is. 
#this means that even though the batch 2 frogs are larger at metamorphosis, that is not the reason for the higher viral loads in Batch 2. 


#3 variable model <-- USING THIS ONE
lm_SQ_all_3 = glm(log(Naive) ~ log(Exposure_duration_numerical) + log(Infected) + Batch, data = Sequential_grouped_nozeroes)
summary(lm_SQ_all_3)
plot(lm_SQ_all_3)

```



Testing for an effect of salinity and heat treatments on viral load: 

All of the animals came out of Nicole's salinity and heat treatments as tadpoles

```{r}
#--------------Visualizing it first
#SHARED
ggplot(data = NoZeroesNaiveSH, aes(x=Larval_condition, y = log(MeanCopyFinal), fill = Larval_condition)) +
  geom_boxplot() +
  scale_fill_brewer(palette = "BrBG") +
  theme(legend.position = "none")
#SEQUENTIAL
ggplot(data = NoZeroesNaiveSQ, aes(x=Larval_condition, y = log(MeanCopyFinal), fill = Larval_condition)) +
  geom_boxplot() +
  scale_fill_brewer(palette = "BrBG") +
  theme(legend.position = "none")

#--------------1-way ANOVA

set.seed(1234)

#SHARED
# Compute the analysis of variance
res_aov_SH <- aov(log(MeanCopyFinal) ~ Larval_condition, data = NoZeroesNaiveSH)
# Summary of the analysis
summary(res_aov_SH)

#SEQUENTIAL
# Compute the analysis of variance
res_aov_SQ <- aov(log(MeanCopyFinal) ~ Larval_condition, data = NoZeroesNaiveSQ)
# Summary of the analysis
summary(res_aov_SQ)



#export a cleaned up csv
write.csv(MetaTransmit, "MetaTransmitCleaned.csv", row.names=FALSE)

```






All the Summary Stats I need 
```{r}

#Summary Stats. 
MetaTransmitNaiveSH$Death_timing <- as.factor(MetaTransmitNaiveSH$Death_timing)
MetaTransmitNaiveSQ$Death_timing <- as.factor(MetaTransmitNaiveSQ$Death_timing)

MetaTransmitNaiveSH$Exposure_duration_numerical <- as.factor(MetaTransmitNaiveSH$Exposure_duration_numerical)

#Death Timing
summary(MetaTransmitNaiveSH$Death_timing)
summary(MetaTransmitNaiveSQ$Death_timing)

#Shared
MetaTransmitNaiveSH$Final_status <- as.factor(MetaTransmitNaiveSH$Final_status)
summary(MetaTransmitNaiveSH$Final_status)

Summary_exposure_status_SH <- MetaTransmitNaiveSH%>% 
  group_by(Exposure_duration_numerical, Final_status_binomial) %>% 
  summarise(n())


Summary_naiveSH <- MetaTransmitNaiveSH%>% 
  group_by(Exposure_duration_numerical, Death_timing) %>% 
  summarise(n())

viral_load_mean_summary_naiveSH <- MetaTransmitNaiveSH%>% 
  group_by(Exposure_duration_numerical) %>% 
  subset(!is.na(MeanCopyFinal)) %>% 
  summarise(mean(MeanCopyFinal))

viral_load_mean_Batch_naiveSH <- MetaTransmitNaiveSH%>% 
  group_by(Batch) %>% 
  subset(!is.na(MeanCopyFinal)) %>% 
  summarise(mean(MeanCopyFinal))

#Sequential
MetaTransmitNaiveSQ$Final_status <- as.factor(MetaTransmitNaiveSQ$Final_status)
MetaTransmitNaiveSQ$Exposure_duration_numerical <- as.factor(MetaTransmitNaiveSQ$Exposure_duration_numerical)
summary(MetaTransmitNaiveSQ$Final_status)
#summarise by exposure period

Summary_exposure_status_SQ <- MetaTransmitNaiveSQ%>% 
  group_by(Exposure_duration_numerical, Final_status_binomial) %>% 
  summarise(n())

#death timing summary
summary(MetaTransmitNaiveSQ$Death_timing)

Summary_naiveSQ <- MetaTransmitNaiveSQ%>% 
  group_by(Exposure_duration_numerical, Death_timing) %>% 
  summarise(n())

Summary_control_pv <- MetaTransmit%>% 
  group_by(Animal_type) %>% 
  summarise(n())

# ------------------
#5 number summaries - a bit clunky but here it is. 

#SEQUENTIAL
simpleSQ <- NoZeroesNaiveSQ %>% 
  select(Exposure_duration, logMeanCopyFinal) %>% 
  mutate(Exposure_duration = as.factor(Exposure_duration))
#1hr
simpleSQ1hr <- simpleSQ %>% 
  subset(Exposure_duration == "1hr") 
summary(simpleSQ1hr$logMeanCopyFinal)
#6hr
simpleSQ6hr <- simpleSQ %>% 
  subset(Exposure_duration == "6hr") 
summary(simpleSQ6hr$logMeanCopyFinal)
#12hr
simpleSQ12hr <- simpleSQ %>% 
  subset(Exposure_duration == "12hr") 
summary(simpleSQ12hr$logMeanCopyFinal)
#24hr
simpleSQ24hr <- simpleSQ %>% 
  subset(Exposure_duration == "24hr") 
summary(simpleSQ24hr$logMeanCopyFinal)
#48hr
simpleSQ48hr <- simpleSQ %>% 
  subset(Exposure_duration == "48hr") 
summary(simpleSQ48hr$logMeanCopyFinal)

#SHARED
#5 number summaries - a bit clunky but here it is. 
simpleSH <- NoZeroesNaiveSH %>% 
  select(Exposure_duration, logMeanCopyFinal) %>% 
  mutate(Exposure_duration = as.factor(Exposure_duration))
#1hr
simpleSH1hr <- simpleSH %>% 
  subset(Exposure_duration == "1hr") 
summary(simpleSH1hr$logMeanCopyFinal)
#6hr
simpleSH6hr <- simpleSH %>% 
  subset(Exposure_duration == "6hr") 
summary(simpleSH6hr$logMeanCopyFinal)
#12hr
simpleSH12hr <- simpleSH %>% 
  subset(Exposure_duration == "12hr") 
summary(simpleSH12hr$logMeanCopyFinal)
#24hr
simpleSH24hr <- simpleSH %>% 
  subset(Exposure_duration == "24hr") 
summary(simpleSH24hr$logMeanCopyFinal)
#48hr
simpleSH48hr <- simpleSH %>% 
  subset(Exposure_duration == "48hr") 
summary(simpleSH48hr$logMeanCopyFinal)

#-------------------

#Citations of packages
citation()
R.version

library(dplyr)
library(report)
report::cite_packages()


library(purrr)

package_bibtex <- c("tidyverse", "ggplot2", "readr", "stringr", "lubridate", "RColorBrewer", "rstatix", "dplyr", "ggpubr", "gridExtra") %>%
  map(citation) %>%
  print(style = "bibtex")

save(package_bibtex, file = "PackageBibtex.BibLaTeX")
```
